---
title: "Plastic Pollution & Economic Development"
subtitle: "A Global Analysis of Corporate Plastic Waste"
author: "Alex Pyatigorsky, Brady O'Shea, and Leo Knudsen"
date: today
format:
  html:
    self-contained: true
    theme: flatly
    toc: true
    toc-depth: 3
    toc-title: "Contents"
    number-sections: true
    code-fold: true
    code-summary: "Show Code"
execute:
  cache: false
  warning: false
  message: false
---

```{r}
#| label: setup
#| include: false

library(tidyverse)
library(lubridate)
library(ggridges)
library(ggtext)
library(scales)
library(broom)
library(knitr)
library(gt)
library(patchwork)
library(ggrepel)
library(readr)

# ── Load Data ──────────────────────────────────────────────────────────────────

plastics <- readr::read_csv(
  'https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2021/2021-01-26/plastics.csv'
)

gdp_df <- read.csv("~/Documents/STAT 431 Projects/Project 1 - Plastic Waste/gdp_data.csv")
names(gdp_df)[names(gdp_df) == "Country"] <- "country_merge"
gdp_df <- gdp_df |> select(-country)
names(gdp_df)[names(gdp_df) == "country_merge"] <- "country"

census_data <- read.csv("~/Downloads/IDB_04-15-2026.csv")
census_data <- census_data[!is.na(census_data$Year), ]
census_data$Year <- as.numeric(census_data$Year)
names(census_data) <- str_to_lower(names(census_data))
names(census_data)[names(census_data) == 'name'] <- 'country'
census_data$population <- as.numeric(gsub(",", "", census_data$population))
census_data$area.in.square.kilometers <- as.numeric(
  gsub(",", "", census_data$area.in.square.kilometers)
)

# ── Harmonize Country Names ────────────────────────────────────────────────────

census_data <- census_data |>
  mutate(country = case_when(
    country == 'Côte d\u2019Ivoire' ~ 'Cote D_ivoire',
    country == 'Korea, South'       ~ 'Korea',
    country == 'Taiwan'             ~ 'Taiwan_ Republic of China (ROC)',
    country == 'United States'      ~ 'United States of America',
    TRUE ~ country
  ))

plastics <- plastics |>
  mutate(country = case_when(
    country == 'ECUADOR' ~ 'Ecuador',
    country == 'NIGERIA' ~ 'Nigeria',
    country == 'United Kingdom of Great Britain & Northern Ireland' ~ 'United Kingdom',
    TRUE ~ country
  ))

# ── Merge Datasets ─────────────────────────────────────────────────────────────

plastics <- merge(plastics, census_data, all.x = TRUE, by = c('country', 'year'))
plastics <- merge(plastics, gdp_df,      all.x = TRUE, by = c('country', 'year'))
```

---

# Data Context

This report draws on three complementary datasets to investigate the relationship
between corporate plastic waste, national population, and economic development.

## Break Free From Plastic Audit Data

The data were collected and published by
[Break Free From Plastic](https://www.breakfreefromplastic.org/), a global
movement working to hold corporations accountable for the plastic pollution their
products generate. Data were collected across dozens of countries worldwide. Each
observation represents a single parent company's plastic waste recorded in a
specific country in a given year, making the country–company–year the
observational unit. The dataset covers two annual brand audit cycles: 2019 and
2020. Volunteer teams organized community cleanup events and manually sorted
collected plastic items by resin type (HDPE, LDPE, PET, PP, PS, PVC, and
"other"), recording the responsible parent company for each piece. The number
of volunteers and cleanup events per country are also included. The audit was
designed to create transparent, brand-level accountability for plastic pollution
and to document which corporations are the largest contributors to the global
plastic crisis.

*Data cleaning note:* The 2019 files contained duplicate polypropylene and
polystyrene columns that were combined and removed. The 2020 grand-total column
was stored as a character type and converted to numeric. Column names were
standardized before merging the two years into one dataset. Three countries were
omitted from per-capita visuals: Taiwan, Montenegro, and unlabeled events
("EMPTY"). Taiwan and Montenegro were removed due to implausibly high plastic
totals relative to their populations, and unlabeled events could not be attributed
to any country.

## U.S. Census Bureau International Database (IDB)

Data were compiled and published by the
[U.S. Census Bureau](https://www.census.gov/programs-surveys/international-programs/about/idb.html)
as part of its International Programs division. The IDB covers countries and
territories worldwide. Each row represents a single country in a single year,
making the country–year the observational unit. The extract used here contains
demographic estimates for 2019 and 2020, matching the plastic audit years. The
Census Bureau derives IDB estimates from a combination of national censuses,
vital registration systems, and demographic surveys, applying demographic
modeling to fill gaps. Population and land area figures from the IDB allow
plastic waste totals to be expressed on a per-capita basis and enable comparisons of waste generation relative to country size.

## World Bank GDP Data

GDP figures were sourced from the
[World Bank Open Data](https://data.worldbank.org/) platform. Coverage is
global, with one observation per country per year. GDP values for 2019 and 2020
were extracted to align with the audit years. The World Bank collects national
accounts data from member governments and international statistical agencies,
reporting both nominal GDP per capita (current USD) and GDP measured at
purchasing-power parity (PPP). GDP per capita provides a measure of national
wealth, enabling analysis of whether economic development is associated with
higher or lower rates of plastic waste generation.

---

# Research Questions

## Question 1 — Wealth and Waste

> **Do wealthier countries produce more plastic waste per capita?**

Countries vary enormously in how much plastic waste they generate, but the
direction of the relationship with national income is not obvious. High-income
countries may produce more consumer packaging, yet they also tend to have more
modernized waste-management technology. This question uses the Break Free From
Plastic audit data joined with World Bank GDP figures to examine whether GDP per capita predicts per-capita plastic waste across countries.

## Question 2 — Population vs. Economic Output

> **Is total plastic waste driven more by a country's population size or its
> economic output?**

Population and GDP per capita are both plausible drivers of plastic waste. A
larger population means more consumers, and higher incomes mean more purchasing
power and more packaged goods. This question uses the IDB census data merged
with the audit and GDP data to investigate these two drivers and assess their
relative importance.

---

# Data Summaries & Visualizations

## Research Question 1: Wealth and Plastic Waste

### Summary Tables

#### Cleanup Activity Statistics

The table below summarizes volunteer participation and cleanup event counts,
produced by iterating a custom summary function across both metrics called summarize_variable.

```{r}
#| label: tbl-cleanup-iter
#| tbl-cap: "Summary statistics for cleanup activity metrics (2019–2020)."

# Custom summary function 
summarize_variable <- function(df, var_name, label) {
  df |>
    rename(.val = all_of(var_name)) |>
    summarise(
      Metric = label,
      Min    = min(.val, na.rm = TRUE),
      Mean   = round(mean(.val, na.rm = TRUE), 1),
      Median = median(.val, na.rm = TRUE),
      SD     = round(sd(.val, na.rm = TRUE), 1),
      IQR    = IQR(.val, na.rm = TRUE),
      Max    = max(.val, na.rm = TRUE)
    )
}

# Iteration across both metrics
map(
  list(
    list(var = "volunteers", label = "Volunteers per Cleanup"),
    list(var = "num_events", label = "Number of Cleanup Events")
  ),
  function(item) summarize_variable(plastics, item$var, item$label)
) |>
  bind_rows() |>
  gt() |>
  tab_header(
    title = "Table 1: Cleanup Activity Summary Statistics",
    subtitle = "Volunteers per cleanup and number of events (2019–2020)"
  ) |>
  cols_label(
    Metric = "Metric",
    Min    = "Min",
    Mean   = "Mean",
    Median = "Median",
    SD     = "Std. Dev.",
    IQR    = "IQR",
    Max    = "Max"
  ) |>
  fmt_number(
    columns = c(Min, Mean, Median, SD, IQR, Max),
    decimals = 1,
    use_seps = TRUE
  ) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) |>
  tab_style(
    style = cell_fill(color = "#2C3E50"),
    locations = cells_column_labels()
  ) |>
  tab_style(
    style = cell_text(color = "white"),
    locations = cells_column_labels()
  ) |>
  tab_options(table.width = pct(80))
```

#### Proportion of Plastic by Resin Type

```{r}
#| label: tbl-plastic-props
#| tbl-cap: "Mean proportion and total count of each plastic resin type across all audits."

# Custom function 
prop_summary <- function(df) {
  df |>
    pivot_longer(
      cols      = c(hdpe, ldpe, o, pet, pp, ps, pvc),
      names_to  = "plastic_type",
      values_to = "count"
    ) |>
    mutate(prop = count / grand_total) |>
    group_by(plastic_type) |>
    summarize(
      mean_prop   = round(mean(prop, na.rm = TRUE), 4),
      total_count = sum(count, na.rm = TRUE)
    )
}

prop_summary(plastics) |>
  gt() |>
  tab_header(
    title = "Table 2: Plastic Waste by Resin Type",
    subtitle = "Mean proportion and total count across all audits"
  ) |>
  cols_label(
    plastic_type = "Plastic Type",
    mean_prop    = "Mean Proportion",
    total_count  = "Total Count"
  ) |>
  fmt_percent(
    columns  = mean_prop,
    decimals = 2
  ) |>
  fmt_integer(
    columns  = total_count,
    use_seps = TRUE
  ) |>
  text_transform(
    locations = cells_body(columns = plastic_type),
    fn = toupper
  ) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) |>
  tab_style(
    style = cell_fill(color = "#18BC9C"),
    locations = cells_column_labels()
  ) |>
  tab_style(
    style = cell_text(color = "white"),
    locations = cells_column_labels()
  ) |>
  tab_spanner(
    label   = "Summary Statistics",
    columns = c(mean_prop, total_count)
  ) |>
  tab_options(table.width = pct(70))
```

### Visualizations

#### Top 20 Countries by Total Plastic Waste

@fig-top20 shows the twenty countries with the highest total recorded plastic
waste, colored by average GDP per capita. High-income countries like the United
States and Switzerland appear alongside lower-income nations such as the Philippines
and India, suggesting that wealth alone does not determine total waste output. Population scale and data-collection capacity could also be playing a role.

```{r}
#| label: fig-top20
#| fig-cap: "Top 20 countries by total plastic waste (2019–2020), colored by average GDP per capita."
#| fig-alt: "Horizontal bar chart showing the top 20 countries by total plastic waste. Bars are colored from dark purple (low GDP) to yellow (high GDP) using a plasma color scale. The Philippines, Nigeria, and Switzerland rank among the highest. A color legend on the right shows GDP per capita in dollars."
#| fig-width: 10
#| fig-height: 7

plastics |>
  group_by(country) |>
  summarise(
    total_plastic = sum(grand_total, na.rm = TRUE),
    avg_gdp       = mean(gdp_per_capita_nominal, na.rm = TRUE)
  ) |>
  filter(!is.na(avg_gdp), total_plastic > 0) |>
  slice_max(total_plastic, n = 20) |>
  ggplot(aes(x = total_plastic, y = reorder(country, total_plastic), fill = avg_gdp)) +
  geom_col() +
  geom_text(aes(label = scales::comma(total_plastic)),
            hjust = -0.1, size = 3, color = "grey30") +
  scale_x_continuous(labels = scales::comma,
                     expand = expansion(mult = c(0, 0.18))) +
  scale_fill_viridis_c(option = "plasma", labels = scales::dollar_format()) +
  labs(
    title    = "Top 20 Countries by Total Plastic Waste",
    subtitle = "Colored by average GDP per capita — wealth and waste do not move in lockstep",
    x        = "Total Plastic Waste (piece count)",
    y        = NULL,
    fill     = "GDP per Capita"
  ) +
  theme_minimal(base_size = 12) +
  theme(panel.grid.major.y = element_blank(),
        legend.position    = "right")
```

#### Plastic Waste Distribution by GDP Group

@fig-ridgeline compares the distribution of total plastic waste across countries
grouped into Low (< \$5,000), Middle (\$5,000–\$20,000), and High (> \$20,000)
GDP per capita tiers. All three groups show wide, right-skewed distributions.
High-GDP countries have a heavier right tail, consistent with greater consumer
plastic use, but Low-GDP countries are not far behind which likely reflects a  weaker
waste-management infrastructure rather than less plastic generation.

```{r}
#| label: fig-ridgeline
#| fig-cap: "Ridge density plot of total plastic waste (log scale) by GDP per capita group."
#| fig-alt: "Ridgeline density plot with three rows: Low GDP in yellow, Middle GDP in orange, and High GDP in red. The x-axis shows total plastic waste on a log scale. High-GDP countries show a wider, heavier right tail, but all three groups overlap substantially, indicating that GDP group alone does not cleanly separate plastic waste levels."

plastics |>
  group_by(country) |>
  summarize(
    total_plastic      = sum(grand_total, na.rm = TRUE),
    avg_gdp_per_capita = mean(gdp_per_capita_nominal, na.rm = TRUE)
  ) |>
  filter(!is.na(avg_gdp_per_capita)) |>
  mutate(
    gdp_group = ifelse(avg_gdp_per_capita > 20000, "High GDP",
                ifelse(avg_gdp_per_capita <  5000, "Low GDP", "Middle GDP")),
    gdp_group = factor(gdp_group, levels = c("Low GDP", "Middle GDP", "High GDP"))
  ) |>
  ggplot(aes(x = total_plastic, y = gdp_group, fill = gdp_group)) +
  geom_density_ridges() +
  scale_x_log10(labels = scales::label_comma()) +
  scale_fill_manual(values = c("Low GDP"    = "#d5f5ef",
                               "Middle GDP" = "#2C3E50",
                               "High GDP"   = "#fff3cd")) +
  labs(
    title    = "Plastic Waste Distribution by GDP Group",
    subtitle = "Countries grouped by <span style='color: #d5f5ef;'>**Low**</span>, <span style='color: #2C3E50;'>**Middle**</span>, and <span style='color: #fff3cd;'>**High**</span> GDP per Capita",
    x        = "Total Plastic Waste (Log Scale)",
    y        = NULL
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "none",
        plot.subtitle   = element_markdown())
```

---

## Research Question 2: Population vs. Economic Output

### Summary Tables

#### Per-Capita Plastic Waste — Top 10 Countries

```{r}
#| label: tbl-percap
#| tbl-cap: "Top 10 countries by plastic waste per capita (population > 5 million), with corresponding GDP (PPP)."

plastics_percap <- plastics |>
  filter(
    !str_detect(country, pattern = "Taiwan"),
    !str_detect(country, pattern = "EMPTY"),
    population > 5000000
  ) |>
  group_by(country) |>
  summarize(
    plastic_percap = sum(grand_total, na.rm = TRUE) / mean(population, na.rm = TRUE),
    gdp_ppp        = mean(gdp_ppp, na.rm = TRUE)
  )

top10 <- (plastics_percap |> arrange(desc(plastic_percap)))[1:10, ]

top10 |>
  gt() |>
  tab_header(
    title    = "Table 3: Top 10 Countries by Plastic Waste Per Capita",
    subtitle = "Countries with population > 5 million; Taiwan and Montenegro excluded"
  ) |>
  cols_label(
    country        = "Country",
    plastic_percap = "Plastic Per Capita",
    gdp_ppp        = "GDP per Capita (PPP)"
  ) |>
  fmt_number(
    columns  = plastic_percap,
    decimals = 5
  ) |>
  fmt_currency(
    columns  = gdp_ppp,
    currency = "USD",
    decimals = 0
  ) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) |>
  tab_style(
    style = cell_fill(color = "#2C3E50"),
    locations = cells_column_labels()
  ) |>
  tab_style(
    style = cell_text(color = "white"),
    locations = cells_column_labels()
  ) |>
  tab_style(
    style = cell_fill(color = "#d5f5ef"),
    locations = cells_body(rows = 1)
  ) |>
  tab_options(table.width = pct(70))
```

#### Mean Plastic Waste, Population, and GDP by GDP Group

The table below is produced by iterating across the three GDP groups and
computing group-level summaries for each.

```{r}
#| label: tbl-country-iter
#| tbl-cap: "Mean plastic waste, population, and GDP per capita by GDP group (iterative computation)."

# Iteration across GDP groups 
gdp_groups <- c("Low GDP", "Middle GDP", "High GDP")

country_level <- plastics |>
  group_by(country) |>
  summarize(
    total_plastic = sum(grand_total, na.rm = TRUE),
    avg_pop       = mean(population, na.rm = TRUE),
    avg_gdp       = mean(gdp_per_capita_nominal, na.rm = TRUE)
  ) |>
  filter(!is.na(avg_gdp)) |>
  mutate(
    gdp_group = ifelse(avg_gdp > 20000, "High GDP",
                ifelse(avg_gdp <  5000, "Low GDP", "Middle GDP")),
    gdp_group = factor(gdp_group, levels = gdp_groups)
  )

map(gdp_groups, function(grp) {
  country_level |>
    filter(gdp_group == grp) |>
    summarise(
      `GDP Group`           = grp,
      `N Countries`         = n(),
      `Mean Total Plastic`  = round(mean(total_plastic, na.rm = TRUE), 0),
      `Mean Population`     = round(mean(avg_pop, na.rm = TRUE), 0),
      `Mean GDP per Capita` = round(mean(avg_gdp, na.rm = TRUE), 0)
    )
}) |>
  bind_rows() |>
  gt() |>
  tab_header(
    title    = "Table 4: Country-Level Averages by GDP Group",
    subtitle = "Iterative summary across Low, Middle, and High GDP per capita groups"
  ) |>
  fmt_integer(
    columns  = c(`N Countries`, `Mean Total Plastic`, `Mean Population`),
    use_seps = TRUE
  ) |>
  fmt_currency(
    columns  = `Mean GDP per Capita`,
    currency = "USD",
    decimals = 0
  ) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) |>
  tab_style(
    style = cell_fill(color = "#2C3E50"),
    locations = cells_column_labels()
  ) |>
  tab_style(
    style = cell_text(color = "white"),
    locations = cells_column_labels()
  ) |>
  tab_style(
    style = cell_fill(color = "#fff3cd"),
    locations = cells_body(rows = `GDP Group` == "High GDP")
  ) |>
  tab_style(
    style = cell_fill(color = "#d5f5ef"),
    locations = cells_body(rows = `GDP Group` == "Low GDP")
  ) |>
  tab_spanner(
    label   = "Country Averages",
    columns = c(`Mean Total Plastic`, `Mean Population`, `Mean GDP per Capita`)
  ) |>
  tab_options(table.width = pct(85))
```

### Visualizations

#### Plastic Waste vs. Population (Log Scale)

@fig-scatter shows a strong positive relationship between population and total
plastic waste on a log–log scale. Larger countries consistently generate more
recorded plastic. The dashed trend line confirms the upward slope, and labeled
outliers reveal that some mid-sized countries (e.g., the Philippines) produce
disproportionately high totals relative to their population.

```{r}
#| label: fig-scatter
#| fig-cap: "Scatter plot of total plastic waste vs. average population (both log scale), with labeled notable countries."
#| fig-alt: "Log-log scatter plot with population on the x-axis and total plastic waste on the y-axis. Each point represents one country shown as a blue dot. A dashed grey trend line runs diagonally upward from lower-left to upper-right. The Philippines, India, and the United States are labeled as high-plastic countries. The overall pattern shows a clear positive relationship between population and plastic waste."
#| fig-width: 10
#| fig-height: 6

scatter_data <- plastics |>
  group_by(country) |>
  summarise(
    total_plastic  = sum(grand_total, na.rm = TRUE),
    avg_population = mean(population, na.rm = TRUE)
  ) |>
  filter(!is.na(avg_population), avg_population > 0, total_plastic > 0)

notable <- scatter_data |>
  arrange(desc(total_plastic)) |>
  slice_head(n = 10) |>
  bind_rows(scatter_data |> arrange(desc(avg_population)) |> slice_head(n = 5)) |>
  distinct()

ggplot(scatter_data, aes(x = avg_population, y = total_plastic)) +
  geom_smooth(method = "lm", se = FALSE, color = "grey60", linetype = "dashed") +
  geom_point(color = "dodgerblue", size = 2.5, alpha = 0.7) +
  ggrepel::geom_text_repel(data = notable, aes(label = country),
                           size = 3.2, color = "grey20",
                           min.segment.length = 0, max.overlaps = 20) +
  scale_x_log10(labels = scales::label_comma()) +
  scale_y_log10(labels = scales::label_comma()) +
  labs(
    title    = "Plastic Waste Scales with Population Size",
    subtitle = "Log-log scale; dashed line shows linear trend",
    x        = "Population (log scale)",
    y        = "Total Plastic Waste (log scale)"
  ) +
  theme_minimal(base_size = 12)
```

#### Per-Capita Plastic Waste vs. GDP PPP — Top 10 Countries

@fig-lollipop juxtaposes the top 10 countries by per-capita plastic waste
(left panel) with their respective GDP (PPP) per capita (right panel). The
comparison reveals no clear monotonic relationship: Switzerland ranks high on
both panels, but several other high-plastic countries have comparatively modest
GDPs. This pattern is consistent with the conclusion that population and
collection effort — not wealth alone — drive observed waste totals.

```{r}
#| label: fig-lollipop
#| fig-cap: "Top 10 countries by plastic waste per capita (left) alongside their GDP per capita PPP (right)."
#| fig-alt: "Side-by-side lollipop charts. The left panel shows plastic per capita with steelblue points connected to the axis by grey segments. The right panel shows GDP PPP with tomato-red points connected by dotdash grey segments. Countries are ordered by plastic per capita on both panels. The two panels do not rank identically, illustrating that GDP and per-capita plastic waste are not tightly correlated among the top countries."
#| fig-width: 10
#| fig-height: 5

plastic_percap_lolipop <- top10 |>
  ggplot(aes(x = plastic_percap, y = country)) +
  geom_segment(aes(yend = country, x = 0, xend = plastic_percap), color = "gray50") +
  geom_point(size = 4, color = "steelblue") +
  labs(
    title = "Plastic Waste Per Capita",
    x     = "Plastic Per Capita",
    y     = ""
  ) +
  theme_minimal()

gdp_percap_lolipop <- top10 |>
  ggplot(aes(x = gdp_ppp, y = country)) +
  geom_segment(aes(x = 0, xend = gdp_ppp), color = "grey25", linetype = "dotdash") +
  geom_point(size = 4, color = "tomato") +
  labs(
    title = "GDP per Capita (PPP)",
    x     = "GDP PPP",
    y     = ""
  ) +
  theme_minimal()

plastic_percap_lolipop + gdp_percap_lolipop +
  plot_annotation(title = "Per-Capita Plastic Waste vs. Economic Output — Top 10 Countries")
```

---

# Data Model

## Model: Total Plastic Waste by Resin Type and Country Area

To formally test whether country size (land area) predicts plastic waste after
accounting for resin type, we fit an ordinary least-squares regression of total
plastic waste on plastic type and country area (km²). Plastic type is treated as
a categorical predictor with HDPE as the reference level.

```{r}
#| label: tbl-model
#| tbl-cap: "OLS regression coefficients — total plastic waste predicted by resin type and country area."

plastic_types <- c("hdpe", "ldpe", "o", "pet", "pp", "ps", "pvc")

tot_plastic_types <- function(type) {
  plastics |>
    group_by(country) |>
    summarize(
      plastic_type = as.character(type),
      total        = sum(get(type), na.rm = TRUE),
      area         = mean(area.in.square.kilometers, na.rm = TRUE)
    )
}

model_data  <- map(plastic_types, tot_plastic_types) |> bind_rows()
LinearModel <- lm(total ~ plastic_type + area, data = model_data)

tidy(LinearModel) |>
  mutate(across(where(is.numeric), \(x) round(x, 3))) |>
  gt() |>
  tab_header(
    title    = "Table 5: Linear Model Coefficients",
    subtitle = "Total plastic waste predicted by resin type and country area"
  ) |>
  cols_label(
    term      = "Term",
    estimate  = "Estimate",
    std.error = "Std. Error",
    statistic = "t Statistic",
    p.value   = "p-value"
  ) |>
  fmt_number(
    columns  = c(estimate, std.error, statistic),
    decimals = 3
  ) |>
  fmt_number(
    columns  = p.value,
    decimals = 3
  ) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) |>
  tab_style(
    style = cell_fill(color = "#2C3E50"),
    locations = cells_column_labels()
  ) |>
  tab_style(
    style = cell_text(color = "white"),
    locations = cells_column_labels()
  ) |>
  tab_style(
    style = cell_fill(color = "#fff3cd"),
    locations = cells_body(rows = p.value < 0.05)
  ) |>
  tab_footnote(
    footnote  = "Reference category: HDPE. Model R² = 0.034; F(7, 867) = 3.29, p = 0.002. Highlighted rows are statistically significant (p < 0.05)."
  ) |>
  tab_options(table.width = pct(85))
```

## Model Interpretation

The linear model regressed total plastic waste on plastic resin type and country
land area. Although the model was statistically significant overall
(F = 3.29, p = 0.002), it explained very little of the variance in plastic waste
(R² = 0.034), indicating that these predictors capture only a small fraction of
what drives waste totals.

Using HDPE as the reference category, only the **"other"** plastic type
differed significantly from HDPE — generating approximately 8,578 more units of
waste on average (p < 0.05). The remaining resin types (LDPE, PET, PP, PS, PVC)
did not differ significantly from HDPE (all p > 0.14).

Critically, **country land area was not a significant predictor** of total
plastic waste (p = 0.738). This finding directly addresses Research Question 2:
the physical size of a country does not meaningfully explain how much plastic
waste it generates. Combined with the scatter-plot evidence in @fig-scatter,
this points toward **population and economic activity** — rather than land area —
as the primary structural drivers of plastic waste output.
